Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-03-07 manuscript/supplemental_material.Rnw
use Spalax instead of Nannospalax
fix significantly typo

\newcommand{\smonkey}[0]{\textit{A. fusciceps}\xspace}
\newcommand{\cow}[0]{\textit{B. taurus}\xspace}
\newcommand{\dog}[0]{\textit{C. familiaris}\xspace}
\newcommand{\armadillo}[0]{\textit{D. novemcinctus}\xspace}
\newcommand{\horse}[0]{\textit{E. caballus}\xspace}
\newcommand{\elephant}[0]{\textit{L. africana}\xspace}
\newcommand{\human}[0]{\textit{H. sapiens}\xspace}
\newcommand{\opossum}[0]{\textit{M. domestica}\xspace}
\newcommand{\mouse}[0]{\textit{M. musculus}\xspace}
\newcommand{\galili}[0]{\textit{N. galili}\xspace}
\newcommand{\sheep}[0]{\textit{O. aries}\xspace}
\newcommand{\bonobo}[0]{\textit{P. paniscus}\xspace}
\newcommand{\carmeli}[0]{\textit{S. carmeli}\xspace}
\newcommand{\pig}[0]{\textit{S. scrofa}\xspace}

@@

\begin{table}
  \begin{tabular}{c p{1.5in} c p{2in}}
    \toprule
    Species                       & Collector & Experiment Number & Sample Number(s) \\
    \midrule
    \textit{Ateles fusciceps}     & Kurt Benirschke & & \\
    \textit{Bos taurus}           & Bob Kreft & & \\
    \textit{Canis familiaris}     & Dalen Agnew & & (Bounty Hubler) \\
    \textit{Dasypus novemcinctus} & Richard Truman, Vincent J. Lynch& & \\
    \textit{Equus caballus}       & \citet{Coleman.ea2010:Structuralannotationofequine} & GSM1139275 & SRR850299,  SRR850300 \\
    \textit{Loxodonta africana}   & Kurt Benirschke & & \\
    \textit{Homo sapiens}         & Nandor Gabor, \citet{Hughes.ea2015:Evaluatingintra-andinter-individual} & GSE66622 
                                             & SRR1850945,
                                               SRR1850946, 
                                               SRR1850951, 
                                               SRR1850952, 
                                               SRR1850953, 
                                               SRR1850954, 
                                               SRR1850957, 
                                               SRR1850958, 
                                               SRR1850965, 
                                               SRR1850966, 
                                               SRR1850975, 
                                               SRR1850976, 
                                               SRR1850983, 
                                               SRR1850984, 
                                               SRR1850987, 
                                               SRR1850988, 
                                               SRR1850991, 
                                               SRR1850992, 
                                               SRR1850993, 
                                               SRR1850994, 
                                               SRR1850995, 
                                               SRR1850996, 
                                               SRR1851003, 
                                               SRR1851004, 
                                               SRR1851011, 
                                               SRR1851012 \\
    \textit{Monodelphis domestica} & Karen Smith, Gunter Wagner & & \\
    \textit{Mus musculus}         & Juan Gonzalez, \citet{Wang.ea2011:surveyfornovelimprinted} & GSE29690 
                                     & SRR243594,
                                       SRR243595,
                                       SRR243596,
                                       SRR243597,
                                       SRR243598,
                                       SRR243599,
                                       SRR243600,
                                       SRR243601,
                                       SRR243602,
                                       SRR243586,
                                       SRR243587,
                                       SRR243588,
                                       SRR243589,
                                       SRR243590,
                                       SRR243591,
                                       SRR243592,
                                       SRR243593\\
    \textit{Nannospalax galili}   & Eviator Nevo & GSE68763 & SRR2016467 \\
    \textit{Ovis aries}           & \citet{Archibald.ea2010:sheepgenomereferencesequence} & ERP005642 & ERR489144, ERR489145\\
    \textit{Pan paniscus}         & William Mintz Fields & & (Teco)\\
    \textit{Spalax carmeli}       & Anne Dupressoir, Eviator Nevo & GSE68763 & SRR2016467 \\
    \textit{Sus scrofa}           & \citet{Ahanda.ea2012:Predictionofaltered3-} &  SRP014902 & SRR543893 \\
    \bottomrule
  \end{tabular}
  \caption{Species, experiment and species numbers used in the analyses presented in this paper}
  \label{tab:srr_numbers}
\end{table}

@@

\begin{table}
  \begin{tabular}{c c c c}
    \toprule
    Species                       & Common Name        & Reference & Read Length \\
    \midrule
    \textit{Ateles fusciceps}     & Spider monkey      & \textit{Homo sapiens} & 100 \\
    \textit{Bos taurus}           & Cow                & UMD3.1     & 75 \\
    \textit{Canis familiaris}     & Dog                & CanFam3.1  & 150 \\
    \textit{Dasypus novemcinctus} & 9-Banded Armadillo & DASNOV3    & 100 \\
    \textit{Equus caballus}       & Horse              & EquCab2    & 100 \\ 
    \textit{Loxodonta africana}   & Elephant           & loxAfr3    & 75 \\
    \textit{Homo sapiens}         & Human              & GRCh38     & 75, 150 \\
    \textit{Monodelphis domestica} & Opossum           & BROAD05    & 75 \\
    \textit{Mus musculus}         & Mouse              & GRCm38     & 35,44,150 \\
    \textit{Nannospalax galili}   & Blind mole rat     & \textit{Mus musculus} & 100 \\
    \textit{Ovis aries}           & Sheep              & Oar\_v3.1   & 151 \\
    \textit{Pan paniscus}         & Bonobo             & CHIMP2.1.4 & 150 \\
    \textit{Spalax carmeli}       & Blind mole rat     & \textit{Mus musculus} & 100 \\
    \textit{Sus scrofa}           & Duroc (pig)        & Sscrofa10.2 & 51 \\
    \bottomrule
  \end{tabular}
  \caption{Species names, common names, reference genomes and read 
    lengths for aligned species. An italic species as a reference 
    indicates that assembled reads were annotated using the corresponding
    reference genome as explained in the text.}
  \label{tab:ref_genome_aligned_species}
\end{table}

@@

<<go.analysis.all.species,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    cat(paste0("\\subsubsection{Signficantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 100 \\textit{",
                           capfirst(this.species),"} expressed transcripts}\n\n"))
    this.caption <- paste0("Signficantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 100 \\textit{",capfirst(this.species),"} expressed transcripts")
    this.label <- paste0("tab:go_terms_enriched_",gsub(" ","_",this.species))
    print(xtable(go.results[grepl("100 No Housekeeping",type) & FDR <= 0.05 & species == this.species,
                            list(GO.ID,Term,FDR,Ontology)
                            ][order(FDR)],
                 caption=this.caption,
                 label=this.label,
                 display=c("s","s","s","e","s")
                 ),
          size="\\small",
          booktabs=TRUE,
          tabular.environment="longtable",
          caption.placement="top",
          floating=FALSE)
}
@

@@

<<go.analysis.top.5p.all.species,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    cat(paste0("\\subsubsection{Signficantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 1\\% of \\textit{",
                           capfirst(this.species),"} expressed transcripts}\n\n"))
    this.caption <- paste0("Signficantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 1\\% of \\textit{",
                           capfirst(this.species),"} expressed transcripts.")
    print(xtable(go.results[grepl("Top 1% No Housekeeping",type) & FDR <= 0.05 & species == this.species,
                            list(GO.ID,Term,FDR,Ontology)
                            ][order(FDR)],
                 caption=this.caption,
                 label=paste0("tab:go_terms_enriched_top_one_",gsub(" ","_",this.species)),
                 display=c("s","s","s","e","s")
                 ),
          size="\\small",
          booktabs=TRUE, 
          tabular.environment="longtable",
          caption.placement="top",
          floating=FALSE)
}
@

>>>>

\newcommand{\smonkey}[0]{\textit{A. fusciceps}\xspace}
\newcommand{\cow}[0]{\textit{B. taurus}\xspace}
\newcommand{\dog}[0]{\textit{C. familiaris}\xspace}
\newcommand{\armadillo}[0]{\textit{D. novemcinctus}\xspace}
\newcommand{\horse}[0]{\textit{E. caballus}\xspace}
\newcommand{\elephant}[0]{\textit{L. africana}\xspace}
\newcommand{\human}[0]{\textit{H. sapiens}\xspace}
\newcommand{\opossum}[0]{\textit{M. domestica}\xspace}
\newcommand{\mouse}[0]{\textit{M. musculus}\xspace}
\newcommand{\galili}[0]{\textit{S. galili}\xspace}
\newcommand{\sheep}[0]{\textit{O. aries}\xspace}
\newcommand{\bonobo}[0]{\textit{P. paniscus}\xspace}
\newcommand{\carmeli}[0]{\textit{S. carmeli}\xspace}
\newcommand{\pig}[0]{\textit{S. scrofa}\xspace}

@@

\begin{table}
  \begin{tabular}{c p{1.5in} c p{2in}}
    \toprule
    Species                       & Collector & Experiment Number & Sample Number(s) \\
    \midrule
    \textit{Ateles fusciceps}     & Kurt Benirschke & & \\
    \textit{Bos taurus}           & Bob Kreft & & \\
    \textit{Canis familiaris}     & Dalen Agnew & & (Bounty Hubler) \\
    \textit{Dasypus novemcinctus} & Richard Truman, Vincent J. Lynch& & \\
    \textit{Equus caballus}       & \citet{Coleman.ea2010:Structuralannotationofequine} & GSM1139275 & SRR850299,  SRR850300 \\
    \textit{Loxodonta africana}   & Kurt Benirschke & & \\
    \textit{Homo sapiens}         & Nandor Gabor, \citet{Hughes.ea2015:Evaluatingintra-andinter-individual} & GSE66622 
                                             & SRR1850945,
                                               SRR1850946, 
                                               SRR1850951, 
                                               SRR1850952, 
                                               SRR1850953, 
                                               SRR1850954, 
                                               SRR1850957, 
                                               SRR1850958, 
                                               SRR1850965, 
                                               SRR1850966, 
                                               SRR1850975, 
                                               SRR1850976, 
                                               SRR1850983, 
                                               SRR1850984, 
                                               SRR1850987, 
                                               SRR1850988, 
                                               SRR1850991, 
                                               SRR1850992, 
                                               SRR1850993, 
                                               SRR1850994, 
                                               SRR1850995, 
                                               SRR1850996, 
                                               SRR1851003, 
                                               SRR1851004, 
                                               SRR1851011, 
                                               SRR1851012 \\
    \textit{Monodelphis domestica} & Karen Smith, Gunter Wagner & & \\
    \textit{Mus musculus}         & Juan Gonzalez, \citet{Wang.ea2011:surveyfornovelimprinted} & GSE29690 
                                     & SRR243594,
                                       SRR243595,
                                       SRR243596,
                                       SRR243597,
                                       SRR243598,
                                       SRR243599,
                                       SRR243600,
                                       SRR243601,
                                       SRR243602,
                                       SRR243586,
                                       SRR243587,
                                       SRR243588,
                                       SRR243589,
                                       SRR243590,
                                       SRR243591,
                                       SRR243592,
                                       SRR243593\\
    \textit{Ovis aries}           & \citet{Archibald.ea2010:sheepgenomereferencesequence} & ERP005642 & ERR489144, ERR489145\\
    \textit{Pan paniscus}         & William Mintz Fields & & (Teco)\\
    \textit{Spalax carmeli}       & Anne Dupressoir, Eviator Nevo & GSE68763 & SRR2016467 \\
    \textit{Spalax galili}        & Eviator Nevo                  & GSE68763 & SRR2016467 \\
    \textit{Sus scrofa}           & \citet{Ahanda.ea2012:Predictionofaltered3-} &  SRP014902 & SRR543893 \\
    \bottomrule
  \end{tabular}
  \caption{Species, experiment and species numbers used in the analyses presented in this paper}
  \label{tab:srr_numbers}
\end{table}

@@

\begin{table}
  \begin{tabular}{c c c c}
    \toprule
    Species                       & Common Name        & Reference & Read Length \\
    \midrule
    \textit{Ateles fusciceps}     & Spider monkey      & \textit{Homo sapiens} & 100 \\
    \textit{Bos taurus}           & Cow                & UMD3.1     & 75 \\
    \textit{Canis familiaris}     & Dog                & CanFam3.1  & 150 \\
    \textit{Dasypus novemcinctus} & 9-Banded Armadillo & DASNOV3    & 100 \\
    \textit{Equus caballus}       & Horse              & EquCab2    & 100 \\ 
    \textit{Loxodonta africana}   & Elephant           & loxAfr3    & 75 \\
    \textit{Homo sapiens}         & Human              & GRCh38     & 75, 150 \\
    \textit{Monodelphis domestica} & Opossum           & BROAD05    & 75 \\
    \textit{Mus musculus}         & Mouse              & GRCm38     & 35,44,150 \\
    \textit{Ovis aries}           & Sheep              & Oar\_v3.1   & 151 \\
    \textit{Pan paniscus}         & Bonobo             & CHIMP2.1.4 & 150 \\
    \textit{Spalax carmeli}       & Blind mole rat     & \textit{Mus musculus} & 100 \\
    \textit{Spalax galili}   & Blind mole rat     & \textit{Mus musculus} & 100 \\
    \textit{Sus scrofa}           & Duroc (pig)        & Sscrofa10.2 & 51 \\
    \bottomrule
  \end{tabular}
  \caption{Species names, common names, reference genomes and read 
    lengths for aligned species. An italic species as a reference 
    indicates that assembled reads were annotated using the corresponding
    reference genome as explained in the text.}
  \label{tab:ref_genome_aligned_species}
\end{table}

@@

<<go.analysis.all.species,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    cat(paste0("\\subsubsection{Significantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 100 \\textit{",
                           capfirst(this.species),"} expressed transcripts}\n\n"))
    this.caption <- paste0("Significantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 100 \\textit{",capfirst(this.species),"} expressed transcripts")
    this.label <- paste0("tab:go_terms_enriched_",gsub(" ","_",this.species))
    print(xtable(go.results[grepl("100 No Housekeeping",type) & FDR <= 0.05 & species == this.species,
                            list(GO.ID,Term,FDR,Ontology)
                            ][order(FDR)],
                 caption=this.caption,
                 label=this.label,
                 display=c("s","s","s","e","s")
                 ),
          size="\\small",
          booktabs=TRUE,
          tabular.environment="longtable",
          caption.placement="top",
          floating=FALSE)
}
@

@@

<<go.analysis.top.5p.all.species,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    cat(paste0("\\subsubsection{Significantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 1\\% of \\textit{",
                           capfirst(this.species),"} expressed transcripts}\n\n"))
    this.caption <- paste0("Significantly enriched ($\\mathrm{FDR}\\le0.05$) GO terms in ",
                           " the top 1\\% of \\textit{",
                           capfirst(this.species),"} expressed transcripts.")
    print(xtable(go.results[grepl("Top 1% No Housekeeping",type) & FDR <= 0.05 & species == this.species,
                            list(GO.ID,Term,FDR,Ontology)
                            ][order(FDR)],
                 caption=this.caption,
                 label=paste0("tab:go_terms_enriched_top_one_",gsub(" ","_",this.species)),
                 display=c("s","s","s","e","s")
                 ),
          size="\\small",
          booktabs=TRUE, 
          tabular.environment="longtable",
          caption.placement="top",
          floating=FALSE)
}
@
